---
title: "DIIG Data Challenge"
author: "Lily Li"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo  =  TRUE,
                      warning = FALSE, 
                      message = FALSE)
```

### Load packages & data

```{r load-packages, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
IBM <- read_csv("data.csv")
```

### Guiding Questions

What factors contribute to employee satisfaction levels and what can IBM do to improve satisfaction? Do certain roles have greater employee churn? If so, what factors lead to this churn?

### Questions

* How much would be appropriate for us to branch out from the guiding questions?
  + Since branching out will lead to less focus on the proposed questions.
* How should we define "employee churn"?
  + Churn: The churn rate is usually calculated as the percentage of employees leaving the company over some specified time period.
   - Do we get to define this "period"?
   - Should we just look which role tend to have shorter **YearsAtCurrentRole**?
   
   
Do some background info, IBM/working conditions, shows understanding
Come up with new questions
As a client, what do you think IMB would be interested in knowing


Formatting question:
would you suggest we knit it into a pdf or a webpage? When we submit the rmd file can it be a pdf so it's easier to look at?

You want to push yourself, explore

treat the presentation like you are presenting to business people
Should explain R^2 we should explain it in the context
Ability to give actionable suggestions
Don't just do stats presentations

Questions should be business-related

about 10min presentation

basically we can ask anything that might be helpful to the employer, you would recommend that...
What should 
what role would you apply for if you want to have higher wage. 
Design who do you want to answer the question to.

Far more important to have great visualizations.

Mostly R and Tableau.

