---
title: "DIIG Data Challenge"
author: "Lily Li"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo  =  TRUE,
                      warning = FALSE, 
                      message = FALSE)
```

### Load packages & data

```{r load-packages, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
IBM <- read_csv("data.csv")
```

### Guiding Questions

What factors contribute to employee satisfaction levels and what can IBM do to improve satisfaction? Do certain roles have greater employee churn? If so, what factors lead to this churn?

### Questions

* How much would be appropriate for us to branch out from the guiding questions?
  + Since branching out will lead to less focus on the proposed questions.
* How should we define "employee churn"?
  + Churn: The churn rate is usually calculated as the percentage of employees leaving the company over some specified time period.
   - Do we get to define this "period"?
   - Should we just look which role tend to have shorter **YearsAtCurrentRole**?
